---
title: "stats project"
author: "BokCHOY"
date: "2024-04-25"
output: html_document
---

```{r}

library(haven)
library(sampleSelection)
library(fastDummies)
art<- read_dta("journal.pone.0296906.s001.dta")
```

```{r}
library(tidyverse)
library(broom)

art<- art %>%
 mutate(sold = ifelse(is.na(Hammerprice), 0,1))

art$Theme <- relevel(factor(art$Theme), ref = "L")
art$Type <- relevel(factor(art$Type), ref = "oil")
art$House <- relevel(factor(art$House), ref = "A")

art <- art %>%
  mutate(surface = SizeLengthhöjd * SizeWidthbredd) %>%
  mutate(surface_squared = surface^2)

art$Month<- substr(art$Event, 5,6)
art$December2remember <- ifelse(art$Month == "12", 1,0)
```


```{r}

type_counts <- table(art$Event)
type_percentages <- prop.table(type_counts) * 100
type_percentages <- round(type_percentages, 2)  # rounding to two decimal places
type_summary <- data.frame(Count = type_counts, Percentage = type_percentages) # Combine counts and percentages into a dataframe
print(type_summary)
```


### Figure 1: Descriptive Statistics for Data Set

```{r}
# Number of Paintings
print(paste("Number of Paintings:", nrow(art)))

# Number of artists, they are stored as numbers so we just take the max
num_unique_values <- max(art$Artist)
print(paste("Number of Artists:", num_unique_values))

# Paintings Sold
num_ones <- sum(art$sold == 1)
print(paste("Count of Paintings Sold:", num_ones, "|| Percentage of Paintings Sold:", round(num_ones/nrow(art),4)*100))

```
```{r}
# Painting Technique
type_counts <- table(art$Type)
type_percentages <- prop.table(type_counts) * 100
type_percentages <- round(type_percentages, 2)  # rounding to two decimal places
type_summary <- data.frame(Count = type_counts, Percentage = type_percentages) # Combine counts and percentages into a dataframe
print(type_summary)
```

```{r}
# Art House
type_counts <- table(art$House)
type_percentages <- prop.table(type_counts) * 100
type_percentages <- round(type_percentages, 2)  # rounding to two decimal places
type_summary <- data.frame(Count = type_counts, Percentage = type_percentages) # Combine counts and percentages into a dataframe
print(type_summary)
```
```{r}
# Sold in December
num_ones <- sum(art$December2remember == 1)
print(paste("Count of Paintings Sold in December:", num_ones, "|| Percentage of Paintings Sold in December:", round(num_ones/nrow(art),4)*100))
```
NOTE: On the table in the article, the count of December paintings sales is listed as 3328 and 50.69%. After examining our code we believe that the researchers may have swapped the numbers for December between December and Non-December.

### Figure 2: Mean and Standard Deviation for Columns by All, Sold, and Unsold
```{r}
# convert this to NAs and then split
art$Pricereal[art$Pricereal == 0] <- NA
# Create dummy variables for the 'Type' column
art = dummy_cols(art, select_columns = "Type")
art = dummy_cols(art, select_columns = "Theme")
art = dummy_cols(art, select_columns = "Theme2")
art = dummy_cols(art, select_columns = "House")

sold_paintings <- art[art$sold == 1, ]
unsold_paintings <- art[art$sold == 0, ]

# Function to calculate mean and standard deviation for specific columns
calc_stats <- function(df) {
  # Specify columns for which you want to calculate statistics
  columns <- c('Pricereal', 'Unsigned1', 'surface', 'surface_squared')
  
  # Calculate mean and standard deviation for specified columns
  stats <- sapply(df[columns], function(x) ifelse(is.factor(x), mean(as.numeric(x)), mean(x)))
  sds <- sapply(df[columns], function(x) ifelse(is.factor(x), sd(as.numeric(x)), sd(x)))
  
  return(data.frame(Mean = stats, SD = sds))
}

# Calculate statistics for paintings
total_stats = calc_stats(art)
sold_stats = calc_stats(sold_paintings)
unsold_stats = calc_stats(unsold_paintings)


print(total_stats)

```
NOTE: No output for price, please see statistics for sold paintings for hammer price

```{r}
print("Stats for Full Data")
dataset = art
print("Stats for Sold Data")
dataset = sold_paintings
print(paste("Mean - Mixed:", round(mean(dataset$Type_mixed),3), " Standard Dev - Mixed:", round(sd(dataset$Type_mixed),3)))
print(paste("Mean - Oil:", round(mean(dataset$Type_oil),3), " Standard Dev - Oil:", round(sd(dataset$Type_oil),2)))
print(paste("Mean - Pastel:", round(mean(dataset$Type_pastel),3), " Standard Dev - Pastel:", round(sd(dataset$Type_pastel),3)))
print(paste("Mean - Water colour:", round(mean(dataset$'Type_water colour'),3), " Standard Dev - Water Colour:", round(sd(dataset$'Type_water colour'),3)))
print("Theme 1")
print(paste("Mean - L:", round(mean(dataset$Theme_L),3), " Standard Dev - L:", round(sd(dataset$Theme_L),3)))
print(paste("Mean - A:", round(mean(dataset$Theme_A),3), " Standard Dev - A", round(sd(dataset$Theme_A),2)))
print(paste("Mean - N:", round(mean(dataset$Theme_N),3), " Standard Dev - N:", round(sd(dataset$Theme_N),3)))
print(paste("Mean - O:", round(mean(dataset$Theme_O),3), " Standard Dev - O:", round(sd(dataset$Theme_O),3)))
print(paste("Mean - P:", round(mean(dataset$Theme_P),3), " Standard Dev - P:", round(sd(dataset$Theme_P),3)))
print(paste("Mean - PO:", round(mean(dataset$Theme_PO),3), " Standard Dev - PO:", round(sd(dataset$Theme_PO),3)))
print(paste("Mean - S:", round(mean(dataset$Theme_S),3), " Standard Dev - S:", round(sd(dataset$Theme_S),3)))
print(paste("Mean - U:", round(mean(dataset$Theme_U),3), " Standard Dev - U:", round(sd(dataset$Theme_U),3)))
print("Theme 2")
print(paste("Mean - 2L:", round(mean(dataset$Theme2_L),3), " Standard Dev - 2L:", round(sd(dataset$Theme2_L),3)))
print(paste("Mean - 2A:", round(mean(dataset$Theme2_A),3), " Standard Dev - 2A", round(sd(dataset$Theme2_A),2)))
print(paste("Mean - 2N:", round(mean(dataset$Theme2_N),3), " Standard Dev - 2N:", round(sd(dataset$Theme2_N),3)))
print(paste("Mean - 2O:", round(mean(dataset$Theme2_O),3), " Standard Dev - 2O:", round(sd(dataset$Theme2_O),3)))
print(paste("Mean - 2P:", round(mean(dataset$Theme2_P),3), " Standard Dev - 2P:", round(sd(dataset$Theme2_P),3)))
print(paste("Mean - 2PO:", round(mean(dataset$Theme2_PO),3), " Standard Dev - 2PO:", round(sd(dataset$Theme2_PO),3)))
print(paste("Mean - 2S:", round(mean(dataset$Theme2_S),3), " Standard Dev - 2S:", round(sd(dataset$Theme2_S),3)))
print(paste("Mean - 2U:", round(mean(dataset$Theme2_U),3), " Standard Dev - 2U:", round(sd(dataset$Theme2_U),3)))
print("Art House")
print(paste("Mean - A:", round(mean(dataset$House_A),3), " Standard Dev - A:", round(sd(dataset$House_A),3)))
print(paste("Mean - B:", round(mean(dataset$House_B),3), " Standard Dev - B", round(sd(dataset$House_B),2)))
print(paste("Mean - U:", round(mean(dataset$House_U),3), " Standard Dev - U:", round(sd(dataset$House_U),3)))

```




```{r}
# Print the statistics for sold paintings
print(sold_stats)
```

```{r}
print("Stats for Sold Data")
dataset = sold_paintings
print(paste("Mean - Mixed:", round(mean(dataset$Type_mixed),3), " Standard Dev - Mixed:", round(sd(dataset$Type_mixed),3)))
print(paste("Mean - Oil:", round(mean(dataset$Type_oil),3), " Standard Dev - Oil:", round(sd(dataset$Type_oil),2)))
print(paste("Mean - Pastel:", round(mean(dataset$Type_pastel),3), " Standard Dev - Pastel:", round(sd(dataset$Type_pastel),3)))
print(paste("Mean - Water colour:", round(mean(dataset$'Type_water colour'),3), " Standard Dev - Water Colour:", round(sd(dataset$'Type_water colour'),3)))
print("Theme 1")
print(paste("Mean - L:", round(mean(dataset$Theme_L),3), " Standard Dev - L:", round(sd(dataset$Theme_L),3)))
print(paste("Mean - A:", round(mean(dataset$Theme_A),3), " Standard Dev - A", round(sd(dataset$Theme_A),2)))
print(paste("Mean - N:", round(mean(dataset$Theme_N),3), " Standard Dev - N:", round(sd(dataset$Theme_N),3)))
print(paste("Mean - O:", round(mean(dataset$Theme_O),3), " Standard Dev - O:", round(sd(dataset$Theme_O),3)))
print(paste("Mean - P:", round(mean(dataset$Theme_P),3), " Standard Dev - P:", round(sd(dataset$Theme_P),3)))
print(paste("Mean - PO:", round(mean(dataset$Theme_PO),3), " Standard Dev - PO:", round(sd(dataset$Theme_PO),3)))
print(paste("Mean - S:", round(mean(dataset$Theme_S),3), " Standard Dev - S:", round(sd(dataset$Theme_S),3)))
print(paste("Mean - U:", round(mean(dataset$Theme_U),3), " Standard Dev - U:", round(sd(dataset$Theme_U),3)))
print("Theme 2")
print(paste("Mean - 2L:", round(mean(dataset$Theme2_L),3), " Standard Dev - 2L:", round(sd(dataset$Theme2_L),3)))
print(paste("Mean - 2A:", round(mean(dataset$Theme2_A),3), " Standard Dev - 2A", round(sd(dataset$Theme2_A),2)))
print(paste("Mean - 2N:", round(mean(dataset$Theme2_N),3), " Standard Dev - 2N:", round(sd(dataset$Theme2_N),3)))
print(paste("Mean - 2O:", round(mean(dataset$Theme2_O),3), " Standard Dev - 2O:", round(sd(dataset$Theme2_O),3)))
print(paste("Mean - 2P:", round(mean(dataset$Theme2_P),3), " Standard Dev - 2P:", round(sd(dataset$Theme2_P),3)))
print(paste("Mean - 2PO:", round(mean(dataset$Theme2_PO),3), " Standard Dev - 2PO:", round(sd(dataset$Theme2_PO),3)))
print(paste("Mean - 2S:", round(mean(dataset$Theme2_S),3), " Standard Dev - 2S:", round(sd(dataset$Theme2_S),3)))
print(paste("Mean - 2U:", round(mean(dataset$Theme2_U),3), " Standard Dev - 2U:", round(sd(dataset$Theme2_U),3)))
print("Art House")
print(paste("Mean - A:", round(mean(dataset$House_A),3), " Standard Dev - A:", round(sd(dataset$House_A),3)))
print(paste("Mean - B:", round(mean(dataset$House_B),3), " Standard Dev - B", round(sd(dataset$House_B),2)))
print(paste("Mean - U:", round(mean(dataset$House_U),3), " Standard Dev - U:", round(sd(dataset$House_U),3)))

```


```{r}
# Print the statistics for unsold paintings
print(unsold_stats)
```

```{r}
print("Stats for Unsold Data")
dataset = unsold_paintings
print(paste("Mean - Mixed:", round(mean(dataset$Type_mixed),3), " Standard Dev - Mixed:", round(sd(dataset$Type_mixed),3)))
print(paste("Mean - Oil:", round(mean(dataset$Type_oil),3), " Standard Dev - Oil:", round(sd(dataset$Type_oil),2)))
print(paste("Mean - Pastel:", round(mean(dataset$Type_pastel),3), " Standard Dev - Pastel:", round(sd(dataset$Type_pastel),3)))
print(paste("Mean - Water colour:", round(mean(dataset$'Type_water colour'),3), " Standard Dev - Water Colour:", round(sd(dataset$'Type_water colour'),3)))
print("Theme 1")
print(paste("Mean - L:", round(mean(dataset$Theme_L),3), " Standard Dev - L:", round(sd(dataset$Theme_L),3)))
print(paste("Mean - A:", round(mean(dataset$Theme_A),3), " Standard Dev - A", round(sd(dataset$Theme_A),2)))
print(paste("Mean - N:", round(mean(dataset$Theme_N),3), " Standard Dev - N:", round(sd(dataset$Theme_N),3)))
print(paste("Mean - O:", round(mean(dataset$Theme_O),3), " Standard Dev - O:", round(sd(dataset$Theme_O),3)))
print(paste("Mean - P:", round(mean(dataset$Theme_P),3), " Standard Dev - P:", round(sd(dataset$Theme_P),3)))
print(paste("Mean - PO:", round(mean(dataset$Theme_PO),3), " Standard Dev - PO:", round(sd(dataset$Theme_PO),3)))
print(paste("Mean - S:", round(mean(dataset$Theme_S),3), " Standard Dev - S:", round(sd(dataset$Theme_S),3)))
print(paste("Mean - U:", round(mean(dataset$Theme_U),3), " Standard Dev - U:", round(sd(dataset$Theme_U),3)))
print("Theme 2")
print(paste("Mean - 2L:", round(mean(dataset$Theme2_L),3), " Standard Dev - 2L:", round(sd(dataset$Theme2_L),3)))
print(paste("Mean - 2A:", round(mean(dataset$Theme2_A),3), " Standard Dev - 2A", round(sd(dataset$Theme2_A),2)))
print(paste("Mean - 2N:", round(mean(dataset$Theme2_N),3), " Standard Dev - 2N:", round(sd(dataset$Theme2_N),3)))
print(paste("Mean - 2O:", round(mean(dataset$Theme2_O),3), " Standard Dev - 2O:", round(sd(dataset$Theme2_O),3)))
print(paste("Mean - 2P:", round(mean(dataset$Theme2_P),3), " Standard Dev - 2P:", round(sd(dataset$Theme2_P),3)))
print(paste("Mean - 2PO:", round(mean(dataset$Theme2_PO),3), " Standard Dev - 2PO:", round(sd(dataset$Theme2_PO),3)))
print(paste("Mean - 2S:", round(mean(dataset$Theme2_S),3), " Standard Dev - 2S:", round(sd(dataset$Theme2_S),3)))
print(paste("Mean - 2U:", round(mean(dataset$Theme2_U),3), " Standard Dev - 2U:", round(sd(dataset$Theme2_U),3)))
print("Art House")
print(paste("Mean - A:", round(mean(dataset$House_A),3), " Standard Dev - A:", round(sd(dataset$House_A),3)))
print(paste("Mean - B:", round(mean(dataset$House_B),3), " Standard Dev - B", round(sd(dataset$House_B),2)))
print(paste("Mean - U:", round(mean(dataset$House_U),3), " Standard Dev - U:", round(sd(dataset$House_U),3)))

```




